\documentclass[11pt]{article}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage{hyperref}
\usepackage{geometry}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{fancyvrb}
\geometry{margin=1in}

\title{Systemics $\Sigma$ --- Minimal Specification}
\author{}
\date{}

\begin{document}
\maketitle

\subsection*{Charter}

\textbf{Charter} \emph{(normative)}

This document gives a domain-agnostic, minimal formal specification of Systemics $\Sigma$. It treats any practice as a kernel-shaped contract that produces decisions from posted evidence under benign variation, with replayable records. No specific domain assumptions (physics, ML, audits, etc.) are required.

\subsection*{Alphabet (Objects \& Maps)}

\textbf{Alphabet} \emph{(normative)}

\begin{itemize}
  \item \(U\): Universe of artifacts.
  \item \(\mathbb{V}\): Valuation space (any measurable space; commonly $\mathbb{R}^k \times \mathbb{B}^m$).
  \item \(\mathbf{2}\): Decision space $\mathbf{2} = \{0,1\}.$
  \item \(\Pi\): Frames (benign contexts).
  \item \(P_n\): Probes (benign perturbations).
  \item \(\Theta\): Floors/thresholds (partially ordered set).
  \item \(\beta\): Invariance budgets (tolerances in a poset/lattice).
  \item \(C\): Capacity budgets (bits/time/energy constraints).
  \item \(\Gamma\): Envelope/meta (versions, seeds, numeric modes, commits).
  \item \(\mathcal{R}\): Records (canonical map $\to$ bytes; hash/ledger optional).
\end{itemize}

\subsection*{Definition: Systemic Kernel}

\textbf{Systemic Kernel} \emph{(normative)}

A systemic kernel is the tuple
\begin{equation}
  K_\mu^\Sigma \;=\; \big(v,\, \chi,\, \Pi,\, P_n,\, \Theta,\, \beta,\, C,\, \Gamma\big),
\end{equation}
where $v:U\to \mathbb{V}$ is a valuation map and $\chi:\mathbb{V}\times\Theta\times\beta\to \mathbf{2}$ is a decision gate.

\subsection*{Metrics \& Order}

\textbf{Wobble and orderings} \emph{(normative)}

\begin{itemize}
  \item A divergence ("wobble") $w:\mathbb{V}\times\mathbb{V}\to \mathbb{R}_{\ge 0}$ on decision-relevant coordinates.
  \item Orders: $\theta\preceq \theta'$ means tightening floors; $\beta'\preceq\beta'$ means loosening invariance budgets; $C'\preceq C$ means reduced capacity.
  \item Metrics: $\rho$ is a metric on the decision space (e.g., $\mathbb{R}_{\ge 0}$ for costs, $\mathbf{2}$ for errors).
  \item A (pseudo-)distance on systemic kernels is $d_\mu(K,K')=\rho(v,v')+\theta(\chi,\chi')+\beta(C,C')$.
\end{itemize}

\subsection*{Axioms}

\textbf{Axiom 1: Existence of Valuation} \emph{(normative)}

$\forall u:U, \exists! v:V, \mathrm{val}(u,v)$.

\textbf{Axiom 2: Universal Decision Gate} \emph{(normative)}

$\forall \theta:\Theta, \exists \chi:\mathbb{V}\times\Theta\times\beta\to \mathbf{2}, \mathrm{gate}(\chi,\theta)$.

\textbf{Axiom 3: Invariance under Reparameterization} \emph{(normative)}

If $\beta' \preceq \beta$ and $\theta' \preceq \theta$, then $\chi'(\cdot,\theta')$ is equivalent to $\chi(\cdot,\theta)$.

\textbf{Axiom 4: Capacity Constraints} \emph{(normative)}

$\forall t, \exists c:C, \mathrm{cap}(t,c)$.

\textbf{Axiom 5: Envelope Uniqueness} \emph{(normative)}

If $\gamma:\mathcal{R}$ is a record, then $\exists! \mu:\Gamma, \mathrm{envelope}(\mu,\gamma)$.

\textbf{Axiom 6: Record-keeping} \emph{(normative)}

$\forall r:\mathcal{R}, \exists e:U\to \mathbb{V}, \mathrm{record}(r,e)$.

\textbf{Axiom 7: Bounded Wobble} \emph{(normative)}

$\forall v,v':V, w(v,v')<\infty$.

\textbf{Axiom 8: Tightening Floors} \emph{(normative)}

$\forall \theta,\theta':\Theta, \theta\preceq\theta' \Rightarrow \chi(\cdot,\theta')\preceq\chi(\cdot,\theta)$.

\subsection*{Conformance}

\textbf{Conformance} \emph{(normative)}

A systemic kernel $K$ conforms to a contract $C$ if $\forall t, \mathrm{exec}(K,t) \in C$.

\subsection*{Morphisms}

\textbf{Morphism} \emph{(normative)}

A morphism between systemic kernels $K,K'$ is a tuple
\begin{equation}
  f = \big(f_v, f_\chi, f_\beta, f_C, f_\Gamma\big),
\end{equation}
where
\begin{itemize}
  \item $f_v:V\to V'$ is a valuation map,
  \item $f_\chi:\mathbb{V}\times\Theta\times\beta\to \mathbb{V}'\times\Theta'\times\beta'$ is a compatible decision map,
  \item $f_\beta:\beta\to\beta'$ and $f_C:C\to C'$ are invariance and capacity transformations,
  \item $f_\Gamma:\Gamma\to\Gamma'$ is an envelope transformation.
\end{itemize}

\textbf{Equivalence of Morphisms} \emph{(normative)}

Morphisms $f$ and $g$ are equivalent, written $f\approx g$, if
\begin{itemize}
  \item $f_v=g_v$,
  \item $f_\chi$ and $g_\chi$ are equivalent decision maps,
  \item $f_\beta=g_\beta$ and $f_C=g_C$,
  \item $f_\Gamma=g_\Gamma$.
\end{itemize}

\subsection*{Recipe}

\textbf{Systemics $\Sigma$ Recipe} \emph{(normative)}

To specify a Systemics $\Sigma$ instance:
\begin{enumerate}
  \item Select a universe $U$ of artifacts.
  \item Define a valuation map $v:U\to \mathbb{V}$.
  \item Specify decision-relevant coordinates using floors $\theta$ and invariance budgets $\beta$.
  \item Define capacity budgets $C$.
  \item Specify an envelope/meta description $\mu$.
  \item Define a decision gate $\chi$.
\end{enumerate}

\subsection*{Notes}

\textbf{Notes} \emph{(informative)}

\begin{itemize}
  \item See also the separate specification for Σ Composition.
  \item This specification is a work in progress; see GitHub for updates.
\end{itemize}

\subsection*{References}

\begin{itemize}
\item GraphFrame K0 (GF0) (spec://\_kernel/gf/gf0-k1)
\item SpecFrame K1 (spec://\_kernel/spec/specframe-k1)
\item Σ Composition (separate spec) (spec://domains/systemics/sigma-composition-k1)
\end{itemize}

\end{document}
