\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{geometry}
\geometry{margin=1in}
\usepackage{amsmath,amssymb,amsthm,mathtools}
\usepackage{hyperref}
\usepackage{enumitem}
\usepackage{microtype}

\title{Systemics Σ — Composition (K1)}
\author{}
\date{}

\begin{document}
\maketitle

\section*{References}
\begin{itemize}
\item GraphFrame K0 (GF0) (spec://\_kernel/gf/gf0-k1)
\item SpecFrame K1 (spec://\_kernel/spec/specframe-k1)
\item Systemics Σ — Minimal Specification (spec://domains/systemics/sigma-k1)
\end{itemize}

\section{Charter}
\subsection*{Charter}
This spec defines how lawful Σ kernels compose in serial, parallel, and temporal forms, preserving Σ lawfulness via posted adapters/combinators and monotone contract combination.

\section{Serial Composition}
\subsection*{Posted adapter requirement}
For serial composition with K1\textasciicircum{}Σ = (v1, χ1, …) and K2\textasciicircum{}Σ = (v2, χ2, …), any adapter f used to produce the second valuation MUST be posted as part of the record/contract. In particular, if v2 is defined from u and the upstream decision by: v2(u) = f(u, χ1(v1(u), Θ1, β1)), then f MUST be included in posted data so downstream decisions remain records-only.
\subsection*{Monotone contract combination}
The composite kernel K2 ∘ K1 is lawful only if (Σ-A1 … Σ-A7) hold for the composite and if the combined contracts (Θ, β, C) are formed via a monotone operator ⊕ (domain-chosen), i.e. tightening in any component must not rely on hidden slack.

\section{Parallel / Product Composition}
\subsection*{Posted decision combinator}
For parallel/product composition K1 ⊗ K2, both kernels evaluate and decisions are aggregated via a posted combinator comb: 2 × 2 → 2 (e.g., AND/OR). The choice of comb MUST be posted as part of the record/contract.
\subsection*{Monotone floors/budgets}
Floors and budgets for parallel composition MUST be combined monotonically (tightening cannot rely on hidden slack).

\section{Temporal Composition (window → page → book)}
\subsection*{Pages and books}
A page is a lawful record κ for one window. A book is an ordered sequence of pages; optionally, pages may be hash-chained. Chaining preserves replay and does not alter decisions.

\section{Composition Invariants}
\subsection*{Records-only preserved}
Composition MUST preserve Σ-A2 (records-only): any dependency introduced by composition must be posted.
\subsection*{Replay preserved}
Composition MUST preserve reflexive reproducibility and replayability obligations implied by the Σ contract.

\end{document}
